---
title: "Generative art using t-SNE"
author: "JBV"
date: "10/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Creating a few vertical streep points.

```{r}
library(tidyverse)
library(ggplot2)
library(viridisLite)
library(viridis)

setup <- function(groups = 10, n = 500, sd = .05) {
  tibble(
    x = rep(1:groups, each = n) + rnorm(groups*n, sd = sd),
    y = rep(seq(from = 0, to = 10, lenght.out = n), groups) +
      rnorm(groups*n, sd = sd),
    group = rep(letters[1:groups], each = n)
  )
}

setup() %>%
  ggplot(aes(x, y, color = group)) +
  geom_point() +
  scale_color_viridis_d(option = "E", guide = "none")
```

Running the t-SNE algorithm

```{r}
library(Rtsne)

do_tsne <- function(coords, perplexity = 300) {
  tsne_fit <- coords %>%
    select(x, y) %>%
    scale() %>%
    Rtsne(perplexity = perplexity, max_iter = 500, check_duplicates = FALSE)
  
    tsne_fit$Y %>%
      as.data.frame() %>%
      cbind(select(coords, -x, -y))
}

setup() %>% # data coords init
  do_tsne() %>%
  ggplot(aes(V1, V2, color = group)) +
  geom_point() +
  scale_color_viridis_d(option = "E", guide = "none") +
  coord_fixed() + theme_void()
```

Making different output plots | efficiently: defining a funciton {make_tsne_plots()} that takes a few key parameters and returns the corresponding plot as a result.

```{r}
make_tsne_plots <- function(groups = 10, n = 200, sd = 0, perplexity = 5) {
  setup(groups = groups, n = n, sd = sd) %>%
    do_tsne(perplexity) %>%
    ggplot(aes(V1, V2, color = group)) +
    geom_point() +
    scale_color_viridis_d(option = "E", guide = "none") +
    coord_fixed() + theme_void() +
    theme(
      plot.margin = margin(20, 20, 20, 20),
      panel.border = element_rect(color = "black", fill = NA)
    )
}

make_tsne_plots(groups = 10, sd = 0)
```

To influence the amount of intersections == modify the perplexity

```{r}
library(patchwork)

make_tsne_plots(groups = 5, sd = 0, perplexity = 2) |
  make_tsne_plots(groups = 5, sd = 0, perplexity = 50) |
  make_tsne_plots(groups = 5, sd = 0, perplexity = 150)
```

To influence the smoothness of the paths == modify the amount of noise applied

```{r}
make_tsne_plots(groups = 10, sd = 0) |
  make_tsne_plots(groups = 10, sd = 0.05) |
  make_tsne_plots(groups = 10, sd = 0.5)
```

t-SNE depends on the random seed, so if we generate the same plot multiple times we get different results.

```{r}
make_tsne_plots(groups = 10, sd = 0, perplexity = 12) |
  make_tsne_plots(groups = 10, sd = 0, perplexity = 12) |
  make_tsne_plots(groups = 10, sd = 0, perplexity = 12)
```
